---
output: pdf_document
editor_options:
  chunk_output_type: console
---
```{r, eval=TRUE, echo=FALSE, message=FALSE}
SEED <- 123
set.seed(SEED)

knitr::opts_chunk$set(
  echo = FALSE,
  eval=TRUE,
  message=FALSE
)

library(tidyverse)
library(latex2exp)
```

# 1. Explain how these estimators relate to your answers to the Exercise in the previous chapter (Ex-3.6.3).
$$
E(X)=\frac{\alpha}{\alpha + 1}
= 1 - \frac{1}{\alpha + 1} 
\Leftrightarrow 1-E(X) = \frac{1}{\alpha + 1}
\Leftrightarrow \frac{1}{1-E(X)} = \alpha + 1
\Leftrightarrow \frac{1}{1-E(X)}-1 = \alpha
\Leftrightarrow \alpha = \frac{E(X)}{1-E(X)}
$$

The expectation of a sample is the mean of that sample.

$$
\hat{\alpha} =
\frac{\bar{x}}
     {1-\bar{x}}
$$

# 2: Download the dataset `ExBetaSim_1.csv` from the data folder, which contains a simulated sample from this distribution. Use both estimators to estimate $\alpha$.

```{r}
D <- read.csv("ExBetaSim_1.csv")

est <- \(x) {
  mu = (x |> reduce(`+`)) / length(x)
  return (mu / (1-mu))
}

est2 <- \(x) {
  mu = (x |> map_vec(log) |> reduce(`+`)) / length(x)
  return (-1 / mu)
}

alpha_hat <- D$X |> est()
alpha_hat2 <- D$X |> est2()
```

For the data set and estimator 1, the estimated $\hat{\alpha}$ is `r alpha_hat`.
For the data set and estimator 2, the estimated $\tilde{\alpha}$ is `r alpha_hat2`.

# 3. Plot the cdf implied by your estimates, and also show the "empirical cumulative density function" of your data, which you can do in `ggplot2` using `stat_ecdf`.

The red line is the first estimator, and the purple line is the second.

```{r warning=FALSE}
F <- \(alpha) \(x) x^alpha
f <- \(alpha) \(x) alpha*x^(alpha-1)

print(ggplot(data=D)
  +geom_segment(x=0,y=0,xend=-0.5,yend=0,arrow=arrow(length = unit(0.25, "cm")))
  +geom_segment(x=1,y=1,xend=1.5,yend=1,arrow=arrow(length = unit(0.25, "cm")))
  +geom_function(fun = F(alpha_hat), color="red")
  +geom_function(fun = F(alpha_hat2), color="purple")
  +stat_ecdf(aes(x=X))
  +ylim(min=0,max=1)
  +ylab(TeX("$F(X)$"))
  +xlim(min=-0.5,max=1.5)
  +xlab(TeX("$X$"))
  +theme_bw()
)
```

```{r}
alpha <- 0.7
F_inv <- \(x) x^(1/alpha)

S <- 10000
N <- 30
D2 <- data.frame(X=sample(
  (expand_grid(simstep=1:S,i=1:N) |>
    mutate(X = runif(n=N*S,min=0,max=1)) |>
    mutate(X = F_inv(X)))$X,
  replace=TRUE,
  size=N
))

alpha_hat <- D2$X |> est()
alpha_hat2 <- D2$X |> est2()
```

# 4. (Simulation exercise) Fix $\alpha=`r alpha`$. Simulate some properties of these estimators for a sample size of $N=`r N`$. Are the estimators biased? Does one stand out as better than the other? Hint: You can simulate the distribution of $X$ by transforming uniform random numbers. Specifically, if $U\sim\mathrm{U}[0,1]$, then: $X = U^{\frac{1}{\alpha}}$ will have the correct distribution.

The random sample taken from $F$ is:

`r D2$X`

The seed for generating this specific sample was `r SEED`.

This sample has an $\hat{\alpha}$ of $`r alpha_hat`$
This sample has an $\tilde{\alpha}$ of $`r alpha_hat2`$

Here is a plot of the data (the red line is an estimate based on the sample, the blue line is the actual cdf):

```{r, warning=FALSE}
print(ggplot(data=D2)
  +geom_segment(x=0,y=0,xend=-0.5,yend=0,arrow=arrow(length = unit(0.25, "cm")))
  +geom_segment(x=1,y=1,xend=1.5,yend=1,arrow=arrow(length = unit(0.25, "cm")))
  +geom_function(fun = F(alpha_hat), color="red")
  +geom_function(fun = F(alpha_hat2), color="purple")
  +geom_function(fun = F(alpha), color="blue")
  +stat_ecdf(aes(x=X))
  +ylim(min=0,max=1)
  +ylab(TeX("$F(X)$"))
  +xlim(min=-0.5,max=1.5)
  +xlab(TeX("$X$"))
  +theme_bw()
)
```

I don't think either estimator is particularly biased. I have experimented with larger sample sizes -
(which you can also do by changing the N variable in the Rmd file) - and the results
speak for themselves. At 10000 values, the data was a perfect match. Is one more biased than the other
at lower values? I don't think so. I could not find a problem with either for the seeds I tested.
I was able to find a few seeds where the second was significantly better than the first.
However, I had also found a few seeds where the first was closest.

In conclusion: They're both kinda good. If I was forced to pick, I'd go for the second.










