---
output: pdf_document
---
```{r, eval=TRUE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  eval=TRUE,
  message=FALSE
)

library(tidyverse)
```

```{r}
F <- \(alpha)\(x) { if (x <= 0) 0 else if (1 <= x) 1 else x^alpha }
```

# 1. What is the support of $X$, and is $X$ a discrete or continuous random variable?

$X$ is a continuous random variable described by $F$ such that:

$$
F: \mathbb{R} \mapsto \mathbb{R}
$$

$$
\left(
  F(x) \doteq
  \begin{cases}
    0        & \text{if } x \leq 0 \\
    x^\alpha & \text{if } 0 < x < 1 \\
    1        & \text{if } x \geq 1
  \end{cases}
\right)
: \alpha \in \mathbb{R}^{+}
$$

$X$ is clearly continuous. I have a proof of this in the "Extras" section, just for fun.

For any continuous random variable $X$ described by a cdf $F$,
that random variable's support can be described as:
$$
S_X = \left\{x ~|~ 0 < \mathrm{F}(x) < 1\right\}
$$
For our function $F$, that occurs at the interval $(0,1)$.

The way the function is written excludes 0 and 1,
as their probability density must be 0.

# 2. calculate the pdf, $f(x)$

pdf is the derivative of the cdf for a continuous random variable.

$$
\frac{\delta F(x)}{\delta x} =
\begin{cases}
  0                     & \text{if } x \leq 0 \\
  \alpha x^{\alpha - 1} & \text{if } 0 < x < 1 \\
  0                     & \text{if } x \geq 1
\end{cases}
$$

# 3. Calculate $E[X]$ and $V[X]$.

In our case (I'm assuming we're meant to assume $iid$):
$$
E(X) = \int_{0}^{1} x \cdot \alpha x^{\alpha - 1} \delta x
$$
$$
\dots = \int_{0}^{1} \alpha x^\alpha \delta x
$$
$$
\dots = \frac{\alpha \cdot 1^{\alpha + 1}}{\alpha + 1} - \frac{\alpha \cdot 0^{\alpha + 1}}{\alpha + 1}
= \frac{\alpha \cdot 1^{\alpha + 1} - \alpha \cdot 0^{\alpha + 1}}{\alpha + 1}
= \frac{\alpha}{\alpha + 1}
$$

$$
V(X) = \int_{0}^{1} x^2 \cdot \alpha x^{\alpha - 1} \delta x - \frac{\alpha^{2}}{(\alpha + 1)^{2}}
$$
$$
\dots = \int_{0}^{1} \alpha x^{\alpha + 1} \delta x - \frac{\alpha^{2}}{(\alpha + 1)^{2}}
$$
$$
\dots = \frac{\alpha}{\alpha + 2} - \frac{\alpha^{2}}{(\alpha + 1)^{2}}
= \frac{\alpha}{a^3 + 4a^2 + 5a + 2}
$$

# Extra:

Given a random variable $X$ described by a cdf $F$, $X$ is continuous iff
$F$ is continuous.

Continuity, as a property in the reals ($\mathbb{R}$), can be expressed:
$$
f: \mathbb{R} \mapsto \mathbb{R} 
$$
$$
\forall y \in \mathbb{R},
F \text{ is continuous at } y \leftrightarrow
\left(
  \lim_{x\rightarrow y^{+}} f(x) = \lim_{x\rightarrow y^{-}} f(x) = f(y)
\right)
~~(A)
$$

The following is a proof that $F$ is continuous:

***

$$
\forall \alpha \in \mathbb{R},
1^\alpha = 1
~~(B)
$$

$$
\forall \alpha \in \mathbb{R} \smallsetminus \{0\},
0^\alpha = 0
~~(C)
$$

$$
\forall \alpha \in \mathbb{R} \smallsetminus \{0\},
x^\alpha \text{ is continuous.}
~~(D)
$$

***

$$
\lim_{x \rightarrow 1^{+}} F(x) = \lim_{x \rightarrow 1^{+}} 1 = 1 ~~(1)
$$
$$
\lim_{x \rightarrow 1^{-}} F(x) = \lim_{x \rightarrow 1^{-}} x^{\alpha} = 1^{\alpha} = 1 ~~(\text{using }A,B,D) ~~(2)
$$
$$
\lim_{x \rightarrow 1^{+}} F(x) = \lim_{x \rightarrow 1^{-}} F(x) = 1 ~~(\text{using }1,2) ~~(3)
$$
$$
\therefore F(x) \text{ is continuous at } 1 ~~(\text{using }A,3) ~~(I)
$$

***

$$
\lim_{x \rightarrow 0^{-}} F(x) = \lim_{x \rightarrow 0^{-}} 0 = 0 ~~(1)
$$
$$
\lim_{x \rightarrow 0^{+}} F(x) = \lim_{x \rightarrow 0^{+}} x^{\alpha} = 0^{\alpha} = 0 ~~(\text{using }A,C,D) ~~(2)
$$
$$
\lim_{x \rightarrow 0^{+}} F(x) = \lim_{x \rightarrow 0^{-}} F(x) = 0 ~~(\text{using }1,2) ~~(3)
$$
$$
\therefore F(x) \text{ is continuous at } 0 ~~(\text{using }A,3) ~~(II)
$$

***

$$
F(x): x \in (0,1) = x^\alpha ~~(1)
$$
$$
x^\alpha \text{ is continuous.} ~~(\text{using }D) ~~(2)
$$
$$
\therefore F(x) \text{ is continuous over } (0,1) ~~(\text{using }1,2) ~~(III)
$$

***

$$
F(x): x \in (1,\infty) = 1 ~~(1)
$$
$$
1 \text{ is continuous.} ~~(\text{trivially}) ~~(2)
$$
$$
\therefore F(x) \text{ is continuous over } (1,\infty) ~~(\text{using }1,2) ~~(IV)
$$

***

$$
F(x): x \in (-\infty,0) = 0 ~~(1)
$$
$$
0 \text{ is continuous.} ~~(\text{trivially}) ~~(2)
$$
$$
\therefore F(x) \text{ is continuous over } (-\infty,0) ~~(\text{using }1,2) ~~(V)
$$

***

$$
\therefore F \text{ is continuous.} ~~(\text{using }I,II,III,IV,V)
$$

***

$$ \blacksquare $$































